---
title: "RedaC:T-Seq: Analysis example"
author: "Dave Surgill"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Libraries for primary data processing
library('data.table')
library('dplyr')
# Libraries for conversion for tx space
library('rtracklayer')
library('GenomicFeatures')
library('genomation')
library('stringr')
```

```{r eval=T}

####################################
# RedaC:T-Seq: Analysis example
#
# This code demonstrates data analysis on 
# RedaC:T-Seq pileup results
#
# Column order for the data:
####################################

# WT NaBH4
# KO NaBH4 (NAT10-/-)
# WT Untreated control

####################################
# Process all mismatches
####################################
pileup <- fread('sampledata/mpileup_output_chr19_min10_parsed.txt')
nrow(pileup)

# Get mismatch percentages
All.ref <- data.frame(pileup, mis_WT_NaBH4=with(pileup, (Sub_0)/(depth_0)), mis_KO_NaBH4=with(pileup, (Sub_1)/(depth_1)), mis_WT_Ctrl=with(pileup, (Sub_2)/(depth_2)),mis_WT_Ctrl_total=with(pileup, ((depth_2) - (Ref_2))/(depth_2)))

#!
#Get ratio between samples
All.ref <- data.frame(All.ref, WT_KO_NaBH4_ratio=with(All.ref, mis_WT_NaBH4/mis_KO_NaBH4), WT_WT_Ctrl_ratio=with(All.ref, mis_WT_NaBH4/mis_WT_Ctrl), KO_WT_NaBH4_ratio=with(All.ref, mis_KO_NaBH4/mis_WT_NaBH4))

head(All.ref)
dim(All.ref)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Apply some filters
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Depth at least 10 in WT AND KO AND Untreated
All.ref <- All.ref %>% filter(depth_0 >= 10 & depth_1 >= 10 & depth_2 >= 10)
dim(All.ref)

#Filter coverage in Substitute  base for > 4 in either treated sample
All.ref <- All.ref %>% filter(Sub_0 >= 4 | Sub_1 >= 4)
dim(All.ref)

#Filter by mismatch fraction >= 1% in either treated sample
All.ref <- All.ref %>% filter(mis_WT_NaBH4 >= 0.01 | mis_KO_NaBH4 >= 0.01)
dim(All.ref)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Obtain ac4C candidate sites (in balanced fashion (KO and WT))
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

Combined.ref.allmm <- All.ref

#~~~~~~~~~~~~~~~~~~~~~~~~
# Append consistent identifiers
#~~~~~~~~~~~~~~~~~~~~~~~~

Combined.ref.allmm$locID <- paste0(Combined.ref.allmm$chr,":",Combined.ref.allmm$loc)
Combined.ref.allmm$subID <- paste0(Combined.ref.allmm$chr,":",Combined.ref.allmm$loc,":",Combined.ref.allmm$ref,">",Combined.ref.allmm$sub)

#~~~~~~~~~~~~~~~~~~~~~~~~
# Save objects (optional)
#~~~~~~~~~~~~~~~~~~~~~~~~

#dim(Combined.ref.allmm)
#save(Combined.ref.allmm, file="Combined.ref.allmm.Rdata")

# These are optional to save for diagnostics
#save(All.ref.unfiltered, file="Robjects/All.ref.unfiltered.Rdata")
#save(Combined.ref.CT, file="Robjects/Combined.ref.CT_pooled.Rdata")
```

# Diagnostics
## Exploring mismatch profiels before resolving stand 
## or performing any filtering

```{r}

Combined.ref.allmm$genomic_conversion <- paste0(Combined.ref.allmm$ref,">",Combined.ref.allmm$sub)

conv_tab <- table(Combined.ref.allmm$genomic_conversion)
conv_tab[order(conv_tab,decreasing=T)]

```


# Map sites to transcripts   


```{r MapToTx_set1, eval=T}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Put sites into BED format 
# Then re-load via genomation
#~~~~~~~~~~~~~~~~~~~~~~~~~~~

NaBH4sites <- Combined.ref.allmm
mybed <- data.frame(chr = NaBH4sites$chr,start = format(NaBH4sites$loc - 1,scientific=F) ,end = format(NaBH4sites$loc,scientific=F),id = NaBH4sites$subID,score = 1,strand = "*")
mybed$start <- str_trim(mybed$start, side = "both")
mybed$end <- str_trim(mybed$end, side = "both")
write.table(mybed,file="genomicsites.bed",col.names=F,quote=F,row.names=F,sep="\t")
mybed <- genomation::readBed("genomicsites.bed")

#~~~~~~~~~~~~~~~~~~~~~~
# Put transcript GTF into
# compatible objects
#~~~~~~~~~~~~~~~~~~~~~~

mytxcanon <- rtracklayer::import("sampledata/genes_chr19.gtf")
myTxDbcanon <- makeTxDbFromGRanges(mytxcanon)

exon_by_tx_canon <- exonsBy(myTxDbcanon, by="tx", use.names=TRUE)
mapped_exon_canon <- mapToTranscripts(mybed, exon_by_tx_canon,ignore.strand=FALSE)

#~~~~~~~~~~~~~~~~~~
# Tabulate feature lengths
#~~~~~~~~~~~~~~~~~

transcriptInfo <- transcriptLengths(myTxDbcanon, with.cds_len=TRUE,
                  with.utr5_len=TRUE, with.utr3_len=TRUE)


gr <-  mapped_exon_canon       
query.idx <- mcols(mapped_exon_canon)[,1] 
query.ids <- mcols(mybed)[,2]
query.sites <- query.ids[query.idx]
length(query.sites)

df <- data.frame(seqnames=seqnames(gr),
  starts=start(gr)-1,
  ends=end(gr),
  names=c(rep(".", length(gr))),
  scores=c(rep(".", length(gr))),
  strand=strand(gr),
  subID = query.sites)


#~~~~~~~~~~~~~~
# Adding annotation, feature location
#~~~~~~~~~~~~~~
        
df$txlength <- transcriptInfo$tx_len[match(df$seqnames,transcriptInfo$gene_id)]        
df$UTR5length <- transcriptInfo$utr5_len[match(df$seqnames,transcriptInfo$gene_id)]        
df$CDSlength <- transcriptInfo$cds_len[match(df$seqnames,transcriptInfo$gene_id)]        
df$UTR3length <- transcriptInfo$utr3_len[match(df$seqnames,transcriptInfo$gene_id)]        

myfeatures <- NULL
myfeatureposs <- NULL

for (i in 1:nrow(df)) {

        myrow <- df[i,]
        mypos <- myrow$starts
        if (myrow$CDSlength < 1) {
          myfeature = "none"
          myfeatures <- c(myfeatures,myfeature)
          myfeatureposs <- c(myfeatureposs,mypos)
        } else {
					if (mypos <= myrow$UTR5length) { myfeature = "UTR5" }
					if (mypos > myrow$UTR5length & mypos <= (myrow$UTR5length + myrow$CDSlength)) { myfeature = "CDS" }
					if (mypos > (myrow$UTR5length + myrow$CDSlength)) { myfeature = "UTR3" }      
					myfeatures <- c(myfeatures,myfeature)
					if (myfeature == "UTR5") { myfeaturepos <- mypos }        
					if (myfeature == "CDS") { myfeaturepos <- mypos - myrow$UTR5length }        
					if (myfeature == "UTR3") { myfeaturepos <- mypos - myrow$UTR5length - myrow$CDSlength }           
					myfeatureposs <- c(myfeatureposs,myfeaturepos)
				}
}
 
df$feature <- myfeatures
df$feature_pos <- myfeatureposs

table(df$feature)
prop.table(table(df$feature))

df$pos_rel_aTIS <- df$starts - df$UTR5length
df$pos_rel_aTIS[df$feature == "none"] <- NA
df$feature_pos[df$feature == "none"] <- NA

#~~~~~~~~~~~~~~~~
# Correct the strand etc
#~~~~~~~~~~~~~~~~~

NaBH4_join <- left_join(df,NaBH4sites,by="subID")

NaBH4_join$ref_strandcorrect <- NaBH4_join$ref
NaBH4_join$sub_strandcorrect <- NaBH4_join$sub

NaBH4_join$ref_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$ref == "C"] <- "G"
NaBH4_join$ref_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$ref == "G"] <- "C"
NaBH4_join$ref_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$ref == "A"] <- "T"
NaBH4_join$ref_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$ref == "T"] <- "A"

NaBH4_join$sub_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$sub == "C"] <- "G"
NaBH4_join$sub_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$sub == "G"] <- "C"
NaBH4_join$sub_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$sub == "A"] <- "T"
NaBH4_join$sub_strandcorrect[NaBH4_join$strand == "-" & NaBH4_join$sub == "T"] <- "A"

NaBH4_join$conversion <- paste0(NaBH4_join$ref_strandcorrect,">",NaBH4_join$sub_strandcorrect)

# Examine strand proportions 
prop.table(table(NaBH4_join$conversion[NaBH4_join$strand == "+"])) * 100

table(NaBH4_join$conversion)
prop.table(table(NaBH4_join$conversion))

#~~~~~~~~~~~~~~~~~~~~~~~
# Find cases of ambiguous mapping
#~~~~~~~~~~~~~~~~~~~~~~

maptab <- table(NaBH4_join$subID)
ambiguous_map <- names(maptab[maptab > 1])
length(ambiguous_map)
table(NaBH4_join$conversion[NaBH4_join$subID %in% ambiguous_map])

#~~~~~~~~~~~~~~~~~~~~~~~
# Find cases of multiple substitution types
#~~~~~~~~~~~~~~~~~~~~~~

testtab <- unique(NaBH4_join[,c("subID","locID")])
multiple_sub_tab <- table(testtab$locID)
multiple_subs <- names(multiple_sub_tab[multiple_sub_tab > 1])


#~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~
# Filter ambiguous mapping or multiple substitutions
#~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~

NaBH4_join_filter <- NaBH4_join %>% filter(!(subID %in% ambiguous_map))
NaBH4_join_filter <- NaBH4_join_filter %>% filter(!(locID %in% multiple_subs))

# testing filter
table(NaBH4_join$subID %in% ambiguous_map)
table(NaBH4_join$locID %in% multiple_subs)


table(NaBH4_join_filter$conversion)
prop.table(table(NaBH4_join_filter$conversion))
barplot(table(NaBH4_join_filter$conversion),las=2,main="Substitutions in unfiltered sites")



#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Save the results
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
NaBH4_allmm_txmapped <- NaBH4_join_filter
save(NaBH4_allmm_txmapped,file="NaBH4_allmm_txmapped.Rdata")    





```

# Filter and get candidates

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Get data for just 1.25%
#~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~

#~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Get candidates that pass comparison to untreated
#~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~

#`````````````````````````````
# Increase minimum depth for comparison to untreated
# Coverage in Untreated >= 25
#`````````````````````````````

#mycandidates <- NaBH4_join_allmm_pool[NaBH4_join_allmm_pool$depth_2 >= 25 & !(is.na(NaBH4_join_allmm_pool$depth_2)),]

mycandidates <- NaBH4_allmm_txmapped

#~~~~~~~~~~~~~~
# Add more fields for diagnostics
#~~~~~~~~~~~~~~
mycandidates$diffMM <- mycandidates$mis_WT_NaBH4 - mycandidates$mis_KO_NaBH4
mycandidates$mean_NaBH4_depth <- (mycandidates$depth_0 + mycandidates$depth_1) / 2
mycandidates$max_NaBH4_MM <- apply(mycandidates[,c("mis_KO_NaBH4","mis_WT_NaBH4")],1,max)
mycandidates$Perc_max_MM_lost <- (mycandidates$max_NaBH4_MM - mycandidates$mis_WT_Ctrl)/mycandidates$max_NaBH4_MM

#`````````````````````````````
# 90% mismatches lost
# This is balanced for WT/KO - takes the maximum of either
#`````````````````````````````
prop.table(table(mycandidates$Perc_max_MM_lost > 0.90))
mycandidates <- mycandidates[mycandidates$Perc_max_MM_lost >= 0.9,]

#`````````````````````````````
# Total mismatches at this site < 1% in WT untreated
#`````````````````````````````
mycandidates <- mycandidates[mycandidates$mis_WT_Ctrl_total < 0.01 ,]

#~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~
# Final selection for testing 
# Raising threshold to 1.25%
#~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~

# Three sets here:
#1 Prefiltered (before anything is applied)
#2 Balanced (Applying the same threshold to WT and KO)
#3 1.25% in WT (As presented in the manuscript)

#prefiltered_candidates <- mycandidates

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Restrict to minimum 1.25% MM in either treated sample
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

balanced_candidates <- mycandidates[mycandidates$mis_WT_NaBH4 >= 0.0125 | mycandidates$mis_KO_NaBH4 >= 0.0125,]


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Perform FET
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

balanced_candidates$Fisher_WT_KO_BH4 = NA
for(i in 1:nrow(balanced_candidates)){
  balanced_candidates$Fisher_WT_KO_BH4[i] = fisher.test(matrix(c(balanced_candidates$Sub_0[i],balanced_candidates$Ref_0[i],balanced_candidates$Sub_1[i],balanced_candidates$Ref_1[i]), nrow=2))$p.value
}

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# FDR correct
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

balanced_candidates$Fisher_WT_KO_BH4.adj <- p.adjust(balanced_candidates$Fisher_WT_KO_BH4,method="BH")




```


# All mismatch types, unadj p-val, balanced candidates

```{r volcano_CT_balanced_unadj, fig.width=10, fig.height=8}


candidates <- balanced_candidates

#~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Variation - PLotting as unadj p, different cutoff
#~~~~~~~~~~~~~~~~~~~~~~~~~~~

conv_tab <- table(candidates$conversion)
conv_tab[order(conv_tab)]

myconvs <- names(conv_tab)

# Set cutoff

myp <- 0.01
myratio <- 5

op <- par(mfrow = c(3,4))
for (i in 1:length(myconvs)) {
  
  myconv <- myconvs[i]
  myset <- candidates[candidates$conversion %in% myconv,]
  
  
  myx <- log2(myset$WT_KO_NaBH4_ratio)
	myy <- -log10(myset$Fisher_WT_KO_BH4)
	myy[myy >= 10] <- 10
	myx[myx >= 6] <- 6
	myy[myy <= -10] <- -10
	myx[myx <= -6] <- -6

	# Get the sigs
	mysig.wt <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$WT_KO_NaBH4_ratio >= myratio,]
	#mysig.wt.ct <- mysig.wt[mysig.wt$conversion == "C>T",]
	myx.wt <- log2(mysig.wt$WT_KO_NaBH4_ratio)
	myy.wt <- -log10(mysig.wt$Fisher_WT_KO_BH4)

	mysig.ko <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$WT_KO_NaBH4_ratio <= (1/myratio),]
	#mysig.ko <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$KO_WT_NaBH4_ratio >= myratio,]
	#mysig.ko.ct <- mysig.ko[mysig.ko$conversion == "C>T",]
	myx.ko <- log2(mysig.ko$WT_KO_NaBH4_ratio)
	myy.ko <- -log10(mysig.ko$Fisher_WT_KO_BH4)

	myy.wt[myy.wt >= 10] <- 10
	myx.wt[myx.wt >= 6] <- 6
	myy.wt[myy.wt <= -10] <- -10
	myx.wt[myx.wt <= -6] <- -6

	myy.ko[myy.ko >= 10] <- 10
	myx.ko[myx.ko >= 6] <- 6
	myy.ko[myy.ko <= -10] <- -10
	myx.ko[myx.ko <= -6] <- -6  
	#~~~~~~~~~~~~~~~~~~~~~
	# Color coding by sig
	#~~~~~~~~~~~~~~~~~~~~~
	plot(myx,myy,main=paste0(myconv,
	" conversion\n(WT vs NAT10-/-)"),xlab="FC",ylab="-log10(P unadj)",col="grey40",las=1,xlim=c(-6,6),ylim=c(0,10),pch=19)
	points(myx.wt,myy.wt,col="red",pch=19)
	points(myx.ko,myy.ko,col="blue",pch=19)
	abline(v=0,col="red",lwd=1)
	abline(v=log2(myratio),col="red",lwd=1,lty=2)   
	abline(v=log2(1/myratio),col="red",lwd=1,lty=2) 
	abline(h=-log10(myp),col="red",lwd=1)   
	legend("top",c(as.character(length(myx.wt)),as.character(length(myx.ko))),pch=19,col=c("red","blue"))
}
par(op)

```



# C>T, unadj p-val, balanced candidates

```{r volcano_CT_balanced, fig.width=5, fig.height=5}


candidates <- balanced_candidates

#~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Variation - PLotting as unadj p, different cutoff
#~~~~~~~~~~~~~~~~~~~~~~~~~~~



# Set cutoff

myp <- 0.01
myratio <- 5


  myconv <- "C>T"
  myset <- candidates[candidates$conversion %in% myconv,]
  
  
  myx <- log2(myset$WT_KO_NaBH4_ratio)
	myy <- -log10(myset$Fisher_WT_KO_BH4)
	myy[myy >= 10] <- 10
	myx[myx >= 6] <- 6
	myy[myy <= -10] <- -10
	myx[myx <= -6] <- -6

	# Get the sigs
	mysig.wt <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$WT_KO_NaBH4_ratio >= myratio,]
	#mysig.wt.ct <- mysig.wt[mysig.wt$conversion == "C>T",]
	myx.wt <- log2(mysig.wt$WT_KO_NaBH4_ratio)
	myy.wt <- -log10(mysig.wt$Fisher_WT_KO_BH4)

	mysig.ko <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$WT_KO_NaBH4_ratio <= (1/myratio),]
	#mysig.ko <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$KO_WT_NaBH4_ratio >= myratio,]
	#mysig.ko.ct <- mysig.ko[mysig.ko$conversion == "C>T",]
	myx.ko <- log2(mysig.ko$WT_KO_NaBH4_ratio)
	myy.ko <- -log10(mysig.ko$Fisher_WT_KO_BH4)

	myy.wt[myy.wt >= 10] <- 10
	myx.wt[myx.wt >= 6] <- 6
	myy.wt[myy.wt <= -10] <- -10
	myx.wt[myx.wt <= -6] <- -6

	myy.ko[myy.ko >= 10] <- 10
	myx.ko[myx.ko >= 6] <- 6
	myy.ko[myy.ko <= -10] <- -10
	myx.ko[myx.ko <= -6] <- -6  
	#~~~~~~~~~~~~~~~~~~~~~
	# Color coding by sig
	#~~~~~~~~~~~~~~~~~~~~~
	plot(myx,myy,main=paste0(myconv,
	" conversion\n(WT vs NAT10-/-)"),xlab="FC",ylab="-log10(P unadj)",col="grey40",las=1,xlim=c(-6,6),pch=19)
	points(myx.wt,myy.wt,col="red",pch=19)
	points(myx.ko,myy.ko,col="blue",pch=19)
	abline(v=0,col="red",lwd=1)
	abline(v=log2(myratio),col="red",lwd=1,lty=2)   
	abline(v=log2(1/myratio),col="red",lwd=1,lty=2) 
	abline(h=-log10(myp),col="red",lwd=1)   
	legend("top",c(as.character(length(myx.wt)),as.character(length(myx.ko))),pch=19,col=c("red","blue"))

```



# Barplot of N values from the Volcano

```{r balanced_volcano_Ns, fig.width=5, fig.height=5}

myp <- 0.05
myratio <- 5

myset <- balanced_candidates
mysig.wt <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$WT_KO_NaBH4_ratio >= myratio,]
mysig.ko <- myset[myset$Fisher_WT_KO_BH4 <= myp & myset$WT_KO_NaBH4_ratio <= (1/myratio),]


tobar <- rbind(table(factor(mysig.wt$conversion,levels=myconvs)),
                     table(factor(mysig.ko$conversion,levels=myconvs)))
barplot(tobar,beside=T,col=c("red","blue"),las=1,main="Mismatch calls (All MM types)\nUnadjusted p <= 0.01, Ratio >= 5",legend.text = c("Up in WT","Up in NAT10-/-"),ylab="Number of sites",xlab="Mismatch type",las=2)

```


# Session info

```{r}
sessionInfo()
```